Probability

### 2016-04-22 prob.tex
prob.tex "\subsection{Plan}" updated...

We now formalize a \emph{quantum probability space} as follows~\cite{BirkhoffVonNeumann1936,gleason1957,Redhead1987-REDINA,DBLP:journals/corr/abs-0910-2393,Maassen2010}.
We first assume an ambient Hilbert space~$\Hilb$ and define the
set of events $\qevents$ as all \emph{projections} on~$\Hilb$.
Each quantum state $\ket{\psi}\in\Hilb\backslash\left\{ 0\right\} $
induces a probability measure $\mu_{\psi}:\qevents\rightarrow[0,1]$
on the space of events defined for any event $E\in\qevents$ as follows\footnote{Recently, people extend the domain of $\mu_{\psi}$ to all operators~$\mathcal{A}$
on $\Hilb$ and consider $\mu_{\psi}:\mathcal{A}\rightarrow\C$~\cite{Maassen2010,Swart2013}.
When an operator $A\in\mathcal{A}$ is Hermitian, $\mu_{\psi}\left(A\right)$
is the expectation value of $A$. We does not take this approach because
we want to focus only on probability. }: 
\[
\mu_{\psi}(E)=\frac{\ip{\psi}{E\psi}}{\ip{\psi}{\psi}}
\]
Similarly to the classical case, this probability measure must satisfy:
\begin{itemize}
\item $\mu(\mathbb{1})=1$, and 
\item for a collection of pairwise orthogonal $E_i$, we have
  $\mu(\sum_i E_i) = \sum_i \mu(E_i)$.

@@

Several assumptions are woven in the definition of a quantum probability space:
\begin{itemize}
\item the Hilbert space $\Hilb$;
\item the real interval $[0,1]$; 
\item the fact that each state induces a probability measure, i.e., the
Born rule~\cite{Born1984,Mermin2007}; 
\item the fact that every probability measure is induced by a state, i.e.,
Gleason's theorem~\cite{gleason1957,peres1995quantum,Redhead1987-REDINA}. 
\end{itemize}
In the remainder of the paper, we examine each of these assumptions
and consider variations motivated by computation of numerical quantities in a
world with limited resources. In particular, we will consider a variant of the Hilbert
space over finite fields~$\mathbb{F}_{p^{2}}^{d}$~\cite{HansonOrtizSabryEtAl2015,DQT2014,geometry2013}.

For the probability values, we will consider set-valued probability
measures~\cite{Artstein1972,PuriRalescu1983}, in particular $\mathscr{L}_{2}=\left\{ \textrm{impossible},\textrm{possible}\right\} $,
where impossible is a singleton set,$\left\{ 0\right\} $, and possible
is an open interval, $\left(0,\infty\right)$. Surprisingly, some
combinations of space and probability will result in an unique probability
measure. In these cases, there is no need to discuss whether there
is a Born rule, because we do not have enough probability to correspond
to every state.

If there may be more than one probability measure, we will discuss
whether there is a Born rule to generate a probability measure from
a state. When the space is $\mathbb{C}^{d}$, we will try to induced
a Born rule from the conventional Born rule. When the space is $\mathbb{F}_{p^{2}}^{d}$,
there is no natural way to induce a probability measure from a state,
so we will set some conditions for a Born rule~$\tilde{\pi}$: 
\begin{itemize}
\item Given a pure state $\ket{\Psi}\in\ffzd{p^{2}}$, a Born-rule~$\tilde{\pi}$
should give a probability~$\tilde{\pi}_{\Psi}$; 
\item $\ip{\Psi}{\Phi}=0\Leftrightarrow\tilde{\pi}_{\Psi}\left(\ket{\Phi}\right)=\tilde{0}$,
where $\tilde{0}$ is $0$ for $\left[0,1\right]$ and $\tilde{0}$
is impossible for $\mathscr{L}_{2}=\left\{ \textrm{impossible},\textrm{possible}\right\} $. 
\item $\tilde{\pi}_{\Psi}\left(\ket{\Phi}\right)=\tilde{\pi}_{\mathbf{U}\ket{\Psi}}\left(\mathbf{U}\ket{\Phi}\right)\textrm{ ,}$where
$\ket{\Psi},\ket{\Phi}\in\ffzd{p^{2}}$ and $\mathbf{U}$ is any unitary
map, i.e., $\mathbf{U}^{\dagger}\mathbf{U}=\mathbbm{1}$. 
\end{itemize}
Notice that when the space is $\mathbb{C}^{d}$, every Born rule we
consider will satisfy these three conditions.

Finally, if there is a Born rule, we will see whether every probability
measure is induced by a state, and establish Gleason's theorem. Notice
that Gleason's theorem only hold when $d\ge3$ in CQT so that we may
expect the situation for $d\ge3$ and $d=2$ are different.

The results can be summarized in the following table:\\
\begin{tabular}{>{\raggedright}m{0.15\columnwidth}>{\raggedright}m{0.15\columnwidth}>{\raggedright}m{0.15\columnwidth}>{\raggedright}m{0.15\columnwidth}>{\raggedright}m{0.15\columnwidth}>{\raggedright}m{0.15\columnwidth}}
\hline 
State space $\Hilb$ & Probability values & How many probability measure? & Is there a nature Born rule? & How many possible Born rule if there is no nature one? & How many probability measure not come from any possible Born rules?\tabularnewline
\hline 
\hline 
$\C^{2}$ & $\left[0,1\right]$  & $\ge2$ & Yes &  & $\ge1$\tabularnewline
\hline 
$\C^{d}$ for $d\ge3$ & $\left[0,1\right]$  & $\ge2$ & Yes &  & $0$\tabularnewline
\hline 
$\C^{d}$ & $\mathscr{L}_{2}$ & $\ge2$ & Yes &  & $\ge1$\tabularnewline
\hline 
$\ff{3^{2}}^{d}$ for $3\ge d\ge2$ & $\left[0,1\right]$  & $\ge2$ & No & $1$ & $\ge1$\tabularnewline
\hline 
$\ff{p^{2}}^{2}$ for $p\ge7$ & $\left[0,1\right]$  & $\ge2$ & No & $\infty$ & $\ge1$\tabularnewline
\hline 
$\ff{7^{2}}^{3}$ & $\left[0,1\right]$  & $1$ & No & $0$ & \tabularnewline
\hline 
$\ffzd{p^{2}}$for $d\ge3$ except $d=p=3$ & $\left[0,1\right]$  & $\ge1$

(Whether $\ge2$ or $1$?) & No & $0$ & \tabularnewline
\hline 
$\ffzd{p^{2}}$ & $\mathscr{L}_{2}$ & $\ge2$ & No & $1$ & $\ge1$\tabularnewline
\hline 
\end{tabular}

>>>>

We now formalize a \emph{quantum probability space} as follows~\cite{BirkhoffVonNeumann1936,gleason1957,Redhead1987-REDINA,DBLP:journals/corr/abs-0910-2393,Maassen2010}.
We first assume an ambient Hilbert space~$\Hilb$ and define the
set of events $\qevents$ as all \emph{projections} on~$\Hilb$.
Each quantum state $\ket{\psi}\in\Hilb\backslash\left\{ 0\right\} $
induces a probability measure $\mu_{\psi}:\qevents\rightarrow[0,1]$
on the space of events defined for any event $E\in\qevents$ as follows\footnote{Recently, people extend the domain of $\mu_{\psi}$ to all operators~$\mathcal{A}$
on $\Hilb$ and consider $\mu_{\psi}:\mathcal{A}\rightarrow\C$~\cite{Maassen2010,Swart2013}.
When an operator $A\in\mathcal{A}$ is Hermitian, $\mu_{\psi}\left(A\right)$
is the expectation value of $A$. We does not take this approach because
we want to focus only on probability. }: 
\begin{equation}
\mu_{\psi}(E)=\frac{\ip{\psi}{E\psi}}{\ip{\psi}{\psi}}\label{eq:Born}
\end{equation}
Similarly to the classical case, this probability measure must satisfy:
\begin{itemize}
\item $\mu(\mathbb{1})=1$, and 
\item for a collection of pairwise orthogonal $E_i$, we have
  $\mu(\sum_i E_i) = \sum_i \mu(E_i)$.

@@

Several assumptions are woven in the definition of a quantum probability space:
\begin{itemize}
\item the Hilbert space $\Hilb$;
\item the real interval $[0,1]$; 
\item the fact that each state induces a probability measure, i.e., the
Born rule (\ref{eq:Born})~\cite{Born1984,Mermin2007}; 
\item the fact that every probability measure is induced by a state, i.e.,
Gleason's theorem~\cite{gleason1957,peres1995quantum,Redhead1987-REDINA}. 
\end{itemize}
In the remainder of the paper, we examine each of these assumptions
and consider variations motivated by computation of numerical quantities
in a world with limited resources. In particular, we will consider
a variant of the Hilbert space over finite fields~$\mathbb{F}_{p^{2}}^{d}$~\cite{HansonOrtizSabryEtAl2015,DQT2014,geometry2013}.

For the probability values, we will consider set-valued probability
measures~\cite{Artstein1972,PuriRalescu1983}, in particular $\mathscr{L}_{2}=\left\{ \textrm{impossible},\textrm{possible}\right\} $,
where impossible is a singleton set,$\left\{ 0\right\} $, and possible
is an open interval, $\left(0,\infty\right)$.

If there may be more than one probability measure, we will discuss
whether we will keep using the Born rule (\ref{eq:Born}) or there
is another formula $\tilde{\mu}$ such that $\tilde{\mu}_{\psi}$
is a probability measure for all $\ket{\psi}\in\Hilb\backslash\left\{ 0\right\} $.

Then, Gleason's theorem states given a probability measure $\mu$
there is a mixed state $\ket{\psi}$ such that $\mu=\mu_{\psi}$,
i.e., the Born rule is surjective\footnote{If we extend the domain of $\mu_{\psi}$ including the mixed states.}.
For any other formula~$\tilde{\mu}$, we can ask whether $\tilde{\mu}$
is surjective as well.

Obveriously, we don't want arbirarily assign a state~$\ket{\psi}\in\Hilb\backslash\left\{ 0\right\} $
with a proabability measure $\tilde{\mu}_{\psi}$, for example, assigning
every state to the same probability measure. We want $\tilde{\mu}_{\psi}$
satisfying the following properties with some physical meaning:
\begin{itemize}
\item $\ip{\psi}{\phi}=0\Leftrightarrow\tilde{\mu}_{\psi}\left(\proj{\phi}\right)=\tilde{0}$,
where $\tilde{0}$ is $0$ for $\left[0,1\right]$ and $\tilde{0}$
is impossible for $\mathscr{L}_{2}=\left\{ \textrm{impossible},\textrm{possible}\right\} $. 
\item $\tilde{\mu}_{\psi}\left(\proj{\phi}\right)=\tilde{\mu}_{\mathbf{U}\ket{\psi}}\left(\mathbf{U}\proj{\phi}\mathbf{U}^{\dagger}\right)\textrm{ ,}$where
$\ket{\psi},\ket{\phi}$ are states and $\mathbf{U}$ is any unitary
map, i.e., $\mathbf{U}^{\dagger}\mathbf{U}=\mathbb{1}$. 
\end{itemize}
And the results can be summarized in the following table:\\
\center{ %
\begin{tabular}{>{\raggedright}m{0.2\columnwidth}>{\raggedright}m{0.2\columnwidth}>{\raggedright}m{0.2\columnwidth}>{\raggedright}m{0.2\columnwidth}}
\hline 
State space $\Hilb$  & Probability values  & Is there a $\tilde{\mu}$ satisfying the given conditions?  & Is the $\tilde{\mu}$ surjective?\tabularnewline
\hline 
\hline 
$\C^{d}$ for $d\ge3$  & $\left[0,1\right]$  & Yes  & Yes\tabularnewline
\hline 
$\C^{d}$  & $\mathscr{L}_{2}$  & Yes  & No\tabularnewline
\hline 
$\ffzd{p^{2}}$for $d\ge3$ except $d=p=3$  & $\left[0,1\right]$  & No  & \tabularnewline
\hline 
$\ffzd{p^{2}}$  & $\mathscr{L}_{2}$  & Yes & No\tabularnewline
\hline 
\end{tabular}}
